---
title: "Replication"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#rm(list=ls())
```


## Libraries and user-defined functions

```{r, echo=FALSE}
source("CODE/functions.R")
```


```{r}
library(sf)
library(spdep)
library(rdrobust) 
#library(rdd)
library(ggspatial)
library(furrr)

#library(ggrepel)
library(fields)
library(spdep)
library(estimatr)
library(santoku)
library(patchwork)
```


## Load data

The replication package does not contain actual data as these cannot be publicly shared due to legal reasons. 

The attached file `data_replication.RData` contains table which allows user to test our code, however, it will not yield results reported in the paper. 

Procedure used to generate this table follows (original `rdata` variable is a table with the actual dataset):

```{r}
# rdata <- rdata %>%
#   mutate(
#     across(
#       -c(KOD_OBEC:year),
#       sample,
#       size = nrow(rdata),
#       replace = FALSE
#     )
#   )
# 
# save(rdata, file = "data_replication.RData")
```

Load data:

```{r}
#load("data_replication.RData")
```

### Variables 

Variables:

KOD_OBEC = "Municipality ID (CISOB)",
NAZ_OBEC = "Municipality name",
KOD_OKRES = "District ID",
year = "Year",
so2avg94 = "SO2 concentration in 1994",
so2avg00 = "SO2 concentration in 2000",
pop_b = "Population as of January 1st",
emig = "Total emigration as of December 31st",
immig = "Total immigration as of December 31st",
netm = "Total net immigration as of December 31st",
emig_rate = "Emigration rate (%)",
immig_rate = "Immigration rate (%)",
netm_rate = "Net immigration rate (%)",
pop_rate = "Population growth (%)",
benefits_ti = "municipality eligible for benefits (treatment intensity)",
benefits_ti_ea = "municipality eligible for benefits (treatment intensity, based on economic activity)",
benefits = "municipality eligible for benefits (indicator variable)",
age_20_29 = "Share of 20-29 age group (census, observed)",
age_30_39 = "Share of 30-39 age group (census, observed)",
age_40_49 = "Share of 40-49 age group (census, observed)",
age_50_59 = "Share of 50-59 age group (census, observed)",
age_60_69 = "Share of 60-69 age group (census, observed)",
age_70_79 = "Share of 70-79 age group (census, observed)",
age_20_29_nn = "Share of 20-29 age group (census, nearest neighbor)",
age_30_39_nn = "Share of 30-39 age group (census, nearest neighbor)",
age_40_49_nn = "Share of 40-49 age group (census, nearest neighbor)",
age_50_59_nn = "Share of 50-59 age group (census, nearest neighbor)",
age_60_69_nn = "Share of 60-69 age group (census, nearest neighbor)",
age_70_79_nn = "Share of 70-79 age group (census, nearest neighbor)",
age_20_29_lin = "Share of 20-29 age group (census, linear)",
age_30_39_lin = "Share of 30-39 age group (census, linear)",
age_40_49_lin = "Share of 40-49 age group (census, linear)",
age_50_59_lin = "Share of 50-59 age group (census, linear)",
age_60_69_lin = "Share of 60-69 age group (census, linear)",
age_70_79_lin = "Share of 70-79 age group (census, linear)",
age_20_29_czso = "Share of 20-29 age group (census and registry data)",
age_30_39_czso = "Share of 30-39 age group (census and registry data)",
age_40_49_czso = "Share of 40-49 age group (census and registry data)",
age_50_59_czso = "Share of 50-59 age group (census and registry data)",
age_60_69_czso = "Share of 60-69 age group (census and registry data)",
age_70_79_czso = "Share of 70-79 age group (census and registry data)",
edu_low_sh = "Share of primary educated (census)",
edu_sec_sh = "Share of secondary educated (census)",
edu_ter_sh = "Share of tertiary educated (census)",
edu_low_sh_nn = "Share of primary educated (census, nearest neighbor)",
edu_sec_sh_nn = "Share of secondary educated (census, nearest neighbor)",
edu_ter_sh_nn = "Share of tertiary educated (census, nearest neighbor)",
edu_low_sh_lin = "Share of primary educated (census, linear)",
edu_sec_sh_lin = "Share of secondary educated (census, linear)",
edu_ter_sh_lin = "Share of tertiary educated (census, linear)",
ppdist_Komorany = "Distance to a power plant (Komorany, km)",
ppdist_Ledvice = "Distance to a power plant (Ledvice, km)",
ppdist_Pocerady = "Distance to a power plant (Pocerady, km)",
ppdist_Prunerov = "Distance to a power plant (Prunerov, km)",
ppdist_Tusimice = "Distance to a power plant (Tusimice, km)",
ppdist_Litvinov = "Distance to a power plant (Litvinov, km)",
ppprox_Komorany = "Power plant in municipality (Komorany, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ppprox_Ledvice = "Power plant in municipality (Ledvice, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ppprox_Litvinov = "Power plant in municipality (Litvinov, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ppprox_Pocerady = "Power plant in municipality (Pocerady, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ppprox_Prunerov = "Power plant in municipality (Prunerov, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ppprox_Tusimice = "Power plant in municipality (Tusimice, 0 = no, 1 = yes, 2 = neighboring municipalities)",
ger_share = "Share of ethnic Germans in 1930",
PHAdist = "Distance to Prague (country capital)",
CZdist = "Distance to the country border",
altitude = "Altitude (m)",
TRI = "Terrain roughness",
Udos = "Unemployment rate (U/Pop 15-64)",
area_segment = "Area ID",
refpoint_lon = "Municipality reference point: longitude",
refpoint_lat = "Municipality reference point: latitude",
centroid_lon = "Municipality centroid: longitude",
centroid_lat = "Municipality centroid: latitude",
group1_0km = "Number of man-made amenities in the municipality (Group 1)",
group2_0km = "Number of man-made amenities in the municipality (Group 2)",
group3_0km = "Number of man-made amenities in the municipality (Group 3)",
group1_20km = "Number of man-made amenities within 20 km driving distance (Group 1)",
group2_20km = "Number of man-made amenities within 20 km driving distance (Group 2)",
group3_20km = "Number of man-made amenities within 20 km driving distance (Group 3)"


### Estimation sample

```{r}
rdata_mig <- rdata %>% 
  filter(year >= 1983, year <= 2015) %>% 
  filter(year <= 1989 | year >= 2000) %>% 
  filter(so2avg94 < 70) %>% 
  mutate(
    period = case_when(
      year < 1995 ~ "prior",
      year > 1995 ~ "after"
    )
  ) %>% 
  drop_na(emig_rate) %>% 
  filter(pop_b > 0)
```

## Descriptive statistics

### Time-inavriant characteristcs

```{r}
rdata_mig %>% 
  filter(year == 2011) %>% 
  select(KOD_OBEC,so2avg94, period, ger_share, altitude, starts_with("group")) %>% 
  pivot_longer(-c(KOD_OBEC,so2avg94, period)) %>% 
  split(.$name) %>% 
  map(function(x) descstat_lm(so2avg94, period, xdata = x)) %>% 
  bind_rows(.id = "variable") %>% 
  filter(period == "after") %>% 
  mutate(so2avg94 = as.character(so2avg94)) %>% 
  replace_na(list(so2avg94 = "all")) %>% 
  select(-period,-type) %>% 
  pivot_longer(c(estimate,std.error)) %>% 
  pivot_wider(names_from = so2avg94) %>% 
  mutate(
    variable = ifelse((row_number() %/% 2)*2 == row_number(),"",variable)
  ) %>% 
  mutate(
    year = ifelse(variable == "ger_share","1930",""),
    year = ifelse(str_detect(variable,"group"),"1993",year)
    ) %>% 
    mutate(
      variable = case_when(
        variable == "altitude" ~ "Altitude (m)",
        variable == "ger_share" ~ "Share of ethnic Germans (\\%)",
        variable == "group1_0km" ~ "Man-made amenities in the municipality (Education and health, n)",
        variable == "group2_0km" ~ "Man-made amenities in the municipality (Culture and sports, n)",
        variable == "group3_0km" ~ "Man-made amenities in the municipality (Public administration, n)",
        variable == "group1_20km" ~ "Man-made amenities within 20 km driving distance (Education and health, n)",
        variable == "group2_20km" ~ "Man-made amenities within 20 km driving distance (Culture and sports, n)",
        variable == "group3_20km" ~ "Man-made amenities within 20 km driving distance (Public administration, n)",
        TRUE ~ ""
      )
    ) %>% 
  select(variable,year,all,`30`:`60`)
```
### Characteristics changing over time

```{r}
rdata %>% 
  filter(year %in% c(1980,1991,2001,2011)) %>% 
  mutate(
    period = case_when(
      year < 1995 ~ "prior",
      year > 1995 ~ "after"
    )
  ) %>% 
  filter(so2avg94 < 70) %>%
  select(KOD_OBEC,year,
        starts_with("so2"),
        age_20_29,
        age_30_39,
        age_40_49,
        age_50_59,
        age_60_69,
        age_70_79,
        edu_low_sh,
        edu_sec_sh,
        edu_ter_sh,
        pop_b) %>% 
  mutate(
    so2 = as.character(so2avg94)
    ) %>% 
  pivot_longer(-c(KOD_OBEC,so2,year)) %>% 
  mutate(
    name = str_c(name,"__",year)
  ) %>% 
  select(-year) %>% 
  drop_na() %>% 
  split(.$name) %>% 
  map(function(x) descstat_lm(so2, xdata = x, xformula = as.Formula(value ~ 1))) %>% 
  bind_rows(.id = "variable") %>% 
  filter(type == "mean") %>% 
  separate(variable, c("variable","year"), sep="__") %>% 
  select(-type) %>% 
  replace_na(list(so2 = "all")) %>% 
  pivot_longer(c(estimate,std.error)) %>% 
  pivot_wider(names_from = so2, values_from = value) %>% 
  mutate(
    variable = case_when(
      variable == "so2avg94" ~ "SO2 concentration in 1994",
      variable == "so2avg00" ~ "SO2 concentration in 2000",
      variable == "age_20_29" ~ "Share of 20-29 age group (census, observed)",
      variable == "age_30_39" ~ "Share of 30-39 age group (census, observed)",
      variable == "age_40_49" ~ "Share of 40-49 age group (census, observed)",
      variable == "age_50_59" ~ "Share of 50-59 age group (census, observed)",
      variable == "age_60_69" ~ "Share of 60-69 age group (census, observed)",
      variable == "age_70_79" ~ "Share of 70-79 age group (census, observed)",
      variable == "edu_low_sh" ~ "Share of primary educated (census)",
      variable == "edu_sec_sh" ~ "Share of secondary educated (census)",
      variable == "edu_ter_sh" ~ "Share of tertiary educated (census)",
      variable == "pop_b" ~ "Population as of January 1st",
      TRUE ~ ""
    )
  ) %>% 
  arrange(variable,year,name) %>% 
  mutate(
    year = ifelse((row_number() %/% 2)*2 == row_number(),"",year)
  ) %>% 
  mutate(
    variable = ifelse(year == 1980, variable, "")
  ) %>% 
  select(variable,year,all,`30`:`60`)
```

### Migration

#### Emigration rate
```{r}
dtabs_print <- function(dtb){
  dtb %>% 
  pivot_longer(c(estimate,std.error)) %>% 
  replace_na(list(so2avg94 = "all")) %>% 
  pivot_wider(names_from = so2avg94) %>% 
  replace_na(list(period = "diff")) %>% 
  select(period,all,`30`:`60`) %>% 
  mutate(
    period = ifelse((row_number() %/% 2)*2 == row_number(),"",period)
  ) %>% 
    mutate(
      period = case_when(
        period == "prior" ~ "Pre-desulfurization period (\\%)",
        period == "after" ~ "Post-desulfurization period (\\%)",
        period == "diff" ~ "Difference (p.p.)",
        TRUE ~ ""
      )
    )
}

dtabs <- rdata_mig %>% 
  select(KOD_OBEC,period,so2avg94,ends_with("rate")) %>% 
  pivot_longer(ends_with("rate")) %>% 
  split(.$name) %>% 
  map(function(x) descstat_lm(so2avg94, period, xdata = x))
  
dtabs$emig_rate %>% dtabs_print() #%>% print_latex("FIGURES/tab2_emig_rate.tex")  
```

#### Immigration rate
```{r}
dtabs$immig_rate %>% dtabs_print() #%>% print_latex("FIGURES/tab2_immig_rate.tex")  
```

#### Net migration rate
```{r}
dtabs$netm_rate %>% dtabs_print() #%>% print_latex("FIGURES/tab2_netm_rate.tex")  
```

### Number of municipalities

```{r}
rdata_mig %>% 
  distinct(KOD_OBEC,so2avg94) %>% 
  group_by(so2avg94) %>% 
  summarise(
    Municipalities = n() ,
    .groups = "drop"
  ) %>%
  mutate(
    obs = "Municipalities (n)",
    all = sum(Municipalities)
  ) %>% 
  pivot_wider(names_from = so2avg94, values_from = Municipalities) 
```



# Estimation

## Definition of formulas

```{r}
# Baseline specifications
model_nl <- as.Formula(
  X ~ I(period == "after" & so2avg94 == 40) +
  I(period == "after" & so2avg94 == 50) +
  I(period == "after" & so2avg94 == 60) + 
  year:period:area_segment | KOD_OBEC + I(period == "after") | 0 | KOD_OBEC
)

modell_nl <- list(
  update(model_nl, emig_rate ~ .),
  update(model_nl, immig_rate ~ .),
  update(model_nl, netm_rate ~ .),
  update(model_nl, pop_rate ~ .)
)

model_nl_40p <- as.Formula(
  X ~ I(period == "after" & so2avg94 >= 40) + 
  year:period:area_segment | KOD_OBEC + I(period == "after") | 0 | KOD_OBEC
)


modell_nl_40p <- list(
  update(model_nl_40p, netm_rate ~ .),
  update(model_nl_40p, emig_rate ~ .),
  update(model_nl_40p, immig_rate ~ .),
  update(model_nl_40p, pop_rate ~ .)
)

# Triple DiD

model_benefits <- as.Formula(X ~ I(period == "after" & so2avg94 == 40) + I(period == "after" & 
    so2avg94 == 40 & benefits) + I(period == "after" & benefits) + year:period:area_segment | 
      KOD_OBEC + I(period == "after")  | 0 | KOD_OBEC)

modell_benefits <- list(
  update(model_benefits, emig_rate ~ .),
  update(model_benefits, immig_rate ~ .),
  update(model_benefits, netm_rate ~ .),
  update(model_benefits, pop_rate ~ .)
)

model_benefitsti <- as.Formula(X ~ I(period == "after" & so2avg94 == 40) + I((period == "after" &
    so2avg94 == 40)*benefits_ti_ea) + I((period == "after")*benefits_ti_ea) + year:period:area_segment | 
      KOD_OBEC + I(period == "after")  | 0 | KOD_OBEC)

modell_benefitsti <- list(
  update(model_benefitsti, emig_rate ~ .),
  update(model_benefitsti, immig_rate ~ .),
  update(model_benefitsti, netm_rate ~ .),
  update(model_benefitsti, pop_rate ~ .)
)

```


## DiD (baseline)

```{r}
modell_nl_40p %>% 
  map(felm, data = rdata_mig) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  mutate(
    term = ifelse(stat == "std.error","",term)
  ) %>% 
  select(term, emig_rate, immig_rate, netm_rate)

modell_nl %>% 
  map(felm, data = rdata_mig) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  mutate(
    term = ifelse(stat == "std.error","",term)
  ) %>% 
  select(term, emig_rate, immig_rate, netm_rate)
```

## Robustness tests

### Unemployment
```{r}
rc_unempl <- rdata %>% 
  mutate(
    Udos = ifelse(year < 2001, NA, Udos),
    Udos = ifelse(year <= 1989, 0, Udos)
  ) %>% 
  group_by(KOD_OBEC) %>% 
  arrange(year) %>% 
  mutate(
    Udos_L1 = lag(Udos)
  ) %>% 
  filter(year >= 1983, year <= 2015) %>% 
  filter(year <= 1989 | year >= 2000) %>% 
  filter(so2avg94 < 70) %>% 
  mutate(
    period = case_when(
      year < 1995 ~ "prior",
      year > 1995 ~ "after"
    )
  ) %>% 
  drop_na(emig_rate) %>% 
  filter(pop_b > 0)
  
rc1 <- modell_nl %>% 
  map(
    update, . ~ . + Udos_L1
  ) %>% 
  map(felm, data = rc_unempl) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "U"
  )

```

### Pollution in post-desulfurization period

```{r}
rc2 <- modell_nl %>% 
  map(
    update, . ~ . + I(period == "after" & so2avg00 == 10) + I(period == "after" & 
    so2avg00 == 15) + I(period == "after" & so2avg00 == 20)
  ) %>% 
  map(felm, data = rdata_mig) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "so2"
  )


```

### Education

```{r}
rdata_edu <- rdata_mig %>% select(-starts_with("edu"),-starts_with("age"))

rdata_edu <- rdata %>% 
  select(KOD_OBEC,year,starts_with("edu"),starts_with("age")) %>% 
  mutate(
    age_20_29_nn = ifelse(is.na(age_20_29_czso),age_20_29_nn,age_20_29_czso), 
    age_30_39_nn = ifelse(is.na(age_30_39_czso),age_30_39_nn,age_30_39_czso), 
    age_40_49_nn = ifelse(is.na(age_40_49_czso),age_40_49_nn,age_40_49_czso), 
    age_50_59_nn = ifelse(is.na(age_50_59_czso),age_50_59_nn,age_50_59_czso), 
    age_60_69_nn = ifelse(is.na(age_60_69_czso),age_60_69_nn,age_60_69_czso), 
    age_70_79_nn = ifelse(is.na(age_70_79_czso),age_70_79_nn,age_70_79_czso), 
    age_20_29_lin = ifelse(is.na(age_20_29_czso),age_20_29_lin,age_20_29_czso), 
    age_30_39_lin = ifelse(is.na(age_30_39_czso),age_30_39_lin,age_30_39_czso), 
    age_40_49_lin = ifelse(is.na(age_40_49_czso),age_40_49_lin,age_40_49_czso), 
    age_50_59_lin = ifelse(is.na(age_50_59_czso),age_50_59_lin,age_50_59_czso), 
    age_60_69_lin = ifelse(is.na(age_60_69_czso),age_60_69_lin,age_60_69_czso), 
    age_70_79_lin = ifelse(is.na(age_70_79_czso),age_70_79_lin,age_70_79_czso)
  ) %>% 
  mutate(
    year = year + 1
  ) %>% 
  left_join(rdata_edu,., by = c("KOD_OBEC", "year"))

# "edu_low_sh"     "edu_sec_sh"     "edu_ter_sh"     "edu_low_sh_nn"  "edu_sec_sh_nn"  "edu_ter_sh_nn"  "edu_low_sh_lin"
# [10] "edu_sec_sh_lin" "edu_ter_sh_lin" "age_20_29"      "age_30_39"      "age_40_49"      "age_50_59"      "age_60_69"      "age_70_79"      "age_20_29_nn"  
# [19] "age_30_39_nn"   "age_40_49_nn"   "age_50_59_nn"   "age_60_69_nn"   "age_70_79_nn"   "age_20_29_lin"  "age_30_39_lin"  "age_40_49_lin"  "age_50_59_lin" 
# [28] "age_60_69_lin"  "age_70_79_lin"  "age_20_29_czso" "age_30_39_czso" "age_40_49_czso" "age_50_59_czso" "age_60_69_czso" "age_70_79_czso"
  
rc3 <- modell_nl %>% 
  map(
    update, . ~ . + edu_sec_sh_nn + edu_ter_sh_nn + age_20_29_nn + age_30_39_nn + age_40_49_nn + age_50_59_nn + age_60_69_nn 
  ) %>% 
  map(felm, data = rdata_edu) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "nn"
  )

rc4 <- modell_nl %>% 
  map(
    update, . ~ . + edu_sec_sh_lin + edu_ter_sh_lin + age_20_29_lin + age_30_39_lin + age_40_49_lin + age_50_59_lin + age_60_69_lin
  ) %>% 
  map(felm, data = rdata_edu) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "lin"
  )

```

```{r}
rdata_chom <- rdata %>% 
  filter(year >= 1983, year <= 2015) %>% 
  filter(year <= 1989 | year >= 2000) %>% 
  mutate(
    so2avg94 = ifelse(so2avg94 > 60, 60, so2avg94)
  ) %>% 
  mutate(
    period = case_when(
      year < 1995 ~ "prior",
      year > 1995 ~ "after"
    )
  ) %>% 
  drop_na(emig_rate) %>% 
  filter(pop_b > 0)

rc5 <- modell_nl %>% 
  map(felm, data = rdata_chom) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "chom"
  )



```

```{r}
rdata_bal <- rdata_mig %>% 
  group_by(KOD_OBEC) %>% 
  add_tally() %>% 
  ungroup() %>% 
  filter(n == max(n))

rc6 <- modell_nl %>% 
  map(felm, data = rdata_bal) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "bal"
  )


```

```{r}
rdata_outl <- rdata_mig %>% 
  group_by(year) %>% 
  mutate(
    top_emig = quantile(emig_rate, probs = 0.99, na.rm = TRUE),
    top_immig = quantile(immig_rate, probs = 0.99, na.rm = TRUE),
    top_netm = quantile(netm_rate, probs = 0.99, na.rm = TRUE),
    top_pop = quantile(pop_rate, probs = 0.99, na.rm = TRUE)
  ) %>% 
  ungroup()

rc7 <- list(
rdata_outl %>% 
  filter(emig_rate < top_emig) %>%
  felm(modell_nl[[1]], data = .) %>% 
  felm_column(),
rdata_outl %>% 
  filter(immig_rate < top_immig) %>%
  felm(modell_nl[[2]], data = .) %>% 
  felm_column(),
rdata_outl %>% 
  filter(netm_rate < top_netm) %>%
  felm(modell_nl[[3]], data = .) %>% 
  felm_column(),
rdata_outl %>% 
  filter(pop_rate < top_pop) %>%
  felm(modell_nl[[4]], data = .) %>% 
  felm_column()
) %>% 
  reduce(full_join) %>% 
  pivot_longer(emig_rate:pop_rate) %>% 
  mutate(
    rc = "outl"
  )


```

Export RC tables

```{r}
rctable <- bind_rows(rc1,rc2,rc3,rc4,rc5,rc6,rc7) %>%
  filter(name != "pop_rate") %>% 
  split(.$name) %>%
  map(
    function(x){
      x <- x %>%
        pivot_wider(names_from = rc) %>%
        delete_rows()

      xn <- first(x$name)

      x %>%
        select(-stat, -name) #%>%
        #print_latex(str_c("FIGURES/robust_",xn,".tex"))
    }
  )
```

### RC tests: Emigration rate
```{r}
rctable[[1]] %>% 
  slice(c(1:6,9,10))
```
### RC tests: Immigration rate
```{r}
rctable[[2]] %>% 
  slice(c(1:6,9,10))
```

### RC tests: Net migration rate
```{r}
rctable[[3]] %>% 
  slice(c(1:6,9,10))
```


### Standard errors adjustment
This part of the script needs to be run separately due to conflicting dependencies.

```{r, eval=FALSE}
modell_nl_40p_kelly <- modell_nl_40p %>% 
  map(formula, collapse = TRUE, update = TRUE) %>% 
  map(update, .~ . + 1)

modell_nl_kelly <- modell_nl %>% 
  map(formula, collapse = TRUE, update = TRUE) %>% 
  map(update, .~ . + 1)

modell_nl_kelly <- c(
  modell_nl_40p_kelly[1:3],
  modell_nl_kelly[1:3]
)

data_nl_kelly <- rdata_mig %>% rename(X = refpoint_lon, Y = refpoint_lat) %>% mutate(id = KOD_OBEC)
```

#### Robust (outliers exclusion)

```{r, eval = FALSE}
library(plyr)
library(sf)
library(spdep)
library(rdrobust)
library(ggspatial)
library(furrr)

library(fields)
library(spdep)
library(estimatr)
library(santoku)
library(patchwork)

source("CODE/functions.R")
source("CODE/spatial_se_panel.R")

## Functions

kelly_wrapper <- function(sm, eq, spdata, maxres){
    out <- Spatial_HAC_Panel(eq,
                       spdata,
                       range_search = seq(from=10,to=150,by=10),
                       Smoothness = sm,
                       residual_upper = abs(maxres),
                       residual_lower = -1*abs(maxres),
                       opt_method = "REML")

    print(sm)

    return(out)
}

kelly_wrapper_safely <- safely(kelly_wrapper)

kelly_hac <- function(eq, sm, spdata, maxres = 1000){
    out <- sm %>%
        map(
            kelly_wrapper_safely, eq = eq, spdata = spdata, maxres = maxres
        )

}

eest <- lm(modell_nl_kelly[[4]], data = data_nl_kelly)
eest95 <- abs(eest$residuals) %>% quantile(.95)

emig_kl <- kelly_hac(
    modell_nl_kelly[[4]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly,
    maxres = eest95
)

eest <- lm(modell_nl_kelly[[5]], data = data_nl_kelly)
eest95 <- abs(eest$residuals) %>% quantile(.95)

immig_kl <- kelly_hac(
    modell_nl_kelly[[5]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly,
    maxres = eest95
)

eest <- lm(modell_nl_kelly[[6]], data = data_nl_kelly)
eest95 <- abs(eest$residuals) %>% quantile(.95)

netm_kl <- kelly_hac(
    modell_nl_kelly[[6]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly,
    maxres = eest95
)
```



```{r, eval=FALSE}
mm <- emig_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

ekl <- emig_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

mm <- immig_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

ikl  <- immig_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

mm <- netm_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

nkl <- netm_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

list(ekl,ikl,nkl) %>% 
  reduce(left_join, by = c("term","stat")) %>% 
  select(-stat) 
```


#### All observations

```{r, eval = FALSE}
library(plyr)
library(sf)
library(spdep)
library(rdrobust)
library(ggspatial)
library(furrr)

library(fields)
library(spdep)
library(estimatr)
library(santoku)
library(patchwork)

source("CODE/functions.R")
source("CODE/spatial_se_panel.R")

## Functions

kelly_wrapper <- function(sm, eq, spdata, maxres){
    out <- Spatial_HAC_Panel(eq,
                       spdata,
                       range_search = seq(from=10,to=150,by=10),
                       Smoothness = sm,
                       residual_upper = abs(maxres),
                       residual_lower = -1*abs(maxres),
                       opt_method = "REML")

    print(sm)

    return(out)
}

kelly_wrapper_safely <- safely(kelly_wrapper)

kelly_hac <- function(eq, sm, spdata, maxres = 1000){
    out <- sm %>%
        map(
            kelly_wrapper_safely, eq = eq, spdata = spdata, maxres = 1000
        )

}

emig_kl <- kelly_hac(
    modell_nl_kelly[[4]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly
)

immig_kl <- kelly_hac(
    modell_nl_kelly[[5]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly
)

netm_kl <- kelly_hac(
    modell_nl_kelly[[6]],
    sm = seq(from = 0.1, to = 1, by = 0.1),
    spdata = data_nl_kelly
)
```



```{r, eval=FALSE}
mm <- emig_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

ekl <- emig_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

mm <- immig_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

ikl  <- immig_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

mm <- netm_kl %>% 
  map_dfr(
    function(x) x$result$Spatial_Parameters
  ) %>% 
  mutate(
    id = row_number()
  ) %>% 
  slice_min(loglik) %>% 
  pull(id)

nkl <- netm_kl[[mm]]$result$HAC %>% 
  as_tibble(rownames = "term") %>% 
  filter(str_detect(term,"so2")) %>% 
  mutate(
    hac.p = add_stars(hac.p)
  ) %>% 
  format_number() %>% 
  mutate(
    coef = str_c(coef,hac.p),
    hac.se = str_c("(",hac.se,")")
  ) %>% 
  select(term:hac.se) %>% 
  pivot_longer(-term, names_to = "stat")

list(ekl,ikl,nkl) %>% 
  reduce(left_join, by = c("term","stat")) %>% 
  select(-stat) 
```


## Placebo

```{r}
rdata_placebo <- rdata_mig %>%
  filter(period %in% c("prior")) %>%
  mutate(
    period = case_when(
      year < 1987 ~ "prior",
      year >= 1987 ~ "after"
    )
  )

modell_nl %>% 
  map(felm, data = rdata_placebo) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate) 



modell_nl %>% 
  map(felm, data = rdata_placebo) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate) 

```


## Factors

### Benefits

```{r}
rdata_benefits <- rdata_mig %>% filter(so2avg94 <= 40)
```

#### Indicator variable
```{r}
modell_benefits %>% 
  map(felm, data = rdata_benefits) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  select(term, stat, indi = emig_rate)
```

#### Treatment intensity
```{r}
modell_benefitsti %>% 
  map(felm, data = rdata_benefits) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  select(term, stat, ti = emig_rate)
```

### Social capital
```{r}
m50 <- modell_nl %>% 
  map(felm, data = filter(rdata_mig, ger_share < 50)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  #delete_rows() %>% 
  select(term, stat, e50 = emig_rate)

m50p <- modell_nl %>% 
  map(felm, data = filter(rdata_mig, ger_share >= 50)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  #delete_rows() %>% 
  select(term, stat, e50p = emig_rate)

m75 <- modell_nl %>% 
  map(felm, data = filter(rdata_mig, ger_share < 75)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  #delete_rows() %>% 
  select(term, stat, e75 = emig_rate)

m75p <- modell_nl %>% 
  map(felm, data = filter(rdata_mig, ger_share >= 75)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  #delete_rows() %>% 
  select(term, stat, e75p = emig_rate)

list(m50,m50p,m75,m75p) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(-stat)
```

### Man-made amenities availability

#### Education & health
```{r}
rdata_mma <- rdata_mig %>% 
  select(starts_with("group")) %>% 
  summarise(across(everything(),median)) %>% 
  rename_all(function(x) str_c(x,"_median")) %>% 
  bind_cols(rdata_mig,.)


g1_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group1_0km <= group1_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate)

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group1_0km > group1_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g1_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm"))
```

  
#### Culture & sports
```{r}
g2_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group2_0km <= group2_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate) 

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group2_0km > group2_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join)%>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g2_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm"))
```

#### Public administration facilities & utilities
```{r}
g3_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group3_0km <= group3_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate)

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group3_0km > group3_0km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g2_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm"))
```

#### Education & health (20 km)
```{r}
g1_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group1_20km <= group1_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate)

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group1_20km > group1_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g1_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm"))
```

#### Public administration facilities & utilities (20 km)
```{r}
g2_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group2_20km <= group2_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate) 

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group2_20km > group2_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join)%>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g2_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm"))
```

#### Public administration facilities & utilities (20 km)
```{r}
g3_0 <- modell_nl %>% 
  map(felm, data = filter(rdata_mma, group3_20km <= group3_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(term, emig_rate, immig_rate, netm_rate)

modell_nl %>% 
  map(felm, data = filter(rdata_mma, group3_20km > group3_20km_median)) %>% 
  map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  select(emig_rate_a = emig_rate, immig_rate_a = immig_rate, netm_rate_a = netm_rate) %>% 
  bind_cols(g2_0,.) %>% 
  select(term,starts_with("emig"),starts_with("immig"),starts_with("netm")) %>% 
  print_latex("FIGURES/mma20_g3.tex")
```

## Air pollution and age and education structure

```{r}
model_structure <- as.Formula(
  X ~ I(period == "after" & so2avg94 == 40) +
  I(period == "after" & so2avg94 == 50) +
  I(period == "after" & so2avg94 == 60) + year:period:area_segment | KOD_OBEC + I(period == "after") | 0 | KOD_OBEC
)

model_structure_rc <- as.Formula(
  X ~ I(period == "after" & so2avg94 == 40) +
  I(period == "after" & so2avg94 == 50) +
  I(period == "after" & so2avg94 == 60) | KOD_OBEC + area_time | 0 | KOD_OBEC
)
```


```{r}
data_structure <- rdata %>% 
  filter(year %in% c(1980,1991,2001,2011)) %>% 
  filter(so2avg94 <= 60) %>% 
  mutate(
    area_time = str_c("FE",area_segment,"_",year),
    period = ifelse(year < 1995, "prior", "after")
  )
```

```{r}
names(data_structure)
depv <- c("age_20_29","age_30_39","age_40_49","age_50_59","age_60_69","age_70_79","edu_low_sh","edu_sec_sh","edu_ter_sh")

data_structure %>% 
  select(KOD_OBEC,year,any_of(depv)) %>% 
  pivot_longer(-c(KOD_OBEC,year)) %>%
  drop_na() %>% 
  group_by(name,year) %>% 
  summarise(
    obs = n(),
    .groups = "drop"
  )
  
list(
  update(model_structure, age_20_29 ~ .),
  update(model_structure, age_30_39 ~ .),
  update(model_structure, age_40_49 ~ .),
  update(model_structure, age_50_59 ~ .),
  update(model_structure, age_60_69 ~ .),
  update(model_structure, edu_low_sh ~ .),
  update(model_structure, edu_sec_sh ~ .),
  update(model_structure, edu_ter_sh ~ .)
) %>% 
map(felm, data = data_structure) %>% 
 map(felm_column) %>% 
  reduce(full_join) %>% 
  delete_rows() %>% 
  mutate(
    term = ifelse(stat == "std.error","",term)
  ) 
```

## PTA test
### Emigration rate
```{r}
testd <- rdata_mig %>%
  filter(period == "prior") %>% 
  filter(so2avg94 < 70)

testf <- as.Formula(X ~ I(so2avg94 == 40 & year == 1983) +
  I(so2avg94 == 40 & year == 1984) +
I(so2avg94 == 40 & year == 1985) +
I(so2avg94 == 40 & year == 1986) +
I(so2avg94 == 40 & year == 1987) +
  I(so2avg94 == 40 & year == 1988) +
  I(so2avg94 == 50 & year == 1983) +
  I(so2avg94 == 50 & year == 1984) +
I(so2avg94 == 50 & year == 1985) +
I(so2avg94 == 50 & year == 1986) +
I(so2avg94 == 50 & year == 1987) +
  I(so2avg94 == 50 & year == 1988) +
  I(so2avg94 == 60 & year == 1983) +
  I(so2avg94 == 60 & year == 1984) +
I(so2avg94 == 60 & year == 1985) +
I(so2avg94 == 60 & year == 1986) +
I(so2avg94 == 60 & year == 1987) +
  I(so2avg94 == 60 & year == 1988) | KOD_OBEC | 0 | KOD_OBEC)  # + year:area_segment

felm(update(testf, emig_rate ~ .), data = testd) %>% 
  tidy(conf.int = TRUE, se.type = "cluster") %>% 
  filter(!str_detect(term,"segment")) %>% 
  mutate(
    year = str_extract(term,"\\d{4}") %>% as.double(),
    so2 = str_extract(term,"94 == \\d{2}") %>% str_remove("94 ==")
  ) %>% 
  ggplot(
    aes(x = as.factor(year), group = so2)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = 3,
    lwd = 0.3
  ) +
  geom_linerange(
    aes(ymin = conf.low, ymax = conf.high),
    position = position_dodge(width = 0.3)
  ) +
  geom_point(
    aes(y = estimate, shape = so2),
    position = position_dodge(width = 0.3),
    fill = "white"
  ) +
  scale_shape_discrete(
    #name = "Sulfur dioxide\nconcentration (1994)",
    name = expression(SO[2] * " concentration (1994)"),
    labels = c(
      "40" = expression("40" * mu * "g"/m^3),
      "50" = expression("50" * mu * "g"/m^3),
      "60" = expression("60" * mu * "g"/m^3)
    )
  ) +
  scale_y_continuous(
    name = "Difference from the reference [p.p.]"
  ) +
  theme_classic(base_family = "Times") +
  theme(
    axis.title.x = element_blank()
  )
```

### Immigration rate
```{r}
felm(update(testf, immig_rate ~ .), data = testd) %>% 
  tidy(conf.int = TRUE, se.type = "cluster") %>% 
  filter(!str_detect(term,"segment")) %>% 
  mutate(
    year = str_extract(term,"\\d{4}") %>% as.double(),
    so2 = str_extract(term,"94 == \\d{2}") %>% str_remove("94 ==")
  ) %>% 
  ggplot(
    aes(x = as.factor(year), group = so2)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = 3,
    lwd = 0.3
  ) +
  geom_linerange(
    aes(ymin = conf.low, ymax = conf.high),
    position = position_dodge(width = 0.3)
  ) +
  geom_point(
    aes(y = estimate, shape = so2),
    position = position_dodge(width = 0.3),
    fill = "white"
  ) +
  scale_shape_discrete(
    #name = "Sulfur dioxide\nconcentration (1994)",
    name = expression(SO[2] * " concentration (1994)"),
    labels = c(
      "40" = expression("40" * mu * "g"/m^3),
      "50" = expression("50" * mu * "g"/m^3),
      "60" = expression("60" * mu * "g"/m^3)
    )
  ) +
  scale_y_continuous(
    name = "Difference from the reference [p.p.]"
  ) +
  theme_classic(base_family = "Times") +
  theme(
    axis.title.x = element_blank()
  )
```

### Net migration rate
```{r}
felm(update(testf, netm_rate ~ .), data = testd) %>% 
  tidy(conf.int = TRUE) %>% 
  mutate(
    year = str_extract(term,"\\d{4}") %>% as.double(),
    so2 = str_extract(term,"94 == \\d{2}") %>% str_remove("94 ==")
  ) %>% 
  filter(!str_detect(term,"segment")) %>% 
  ggplot(
    aes(x = as.factor(year), group = so2)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = 3,
    lwd = 0.3
  ) +
  geom_linerange(
    aes(ymin = conf.low, ymax = conf.high),
    position = position_dodge(width = 0.3)
  ) +
  geom_point(
    aes(y = estimate, shape = so2),
    position = position_dodge(width = 0.3),
    fill = "white"
  ) +
  scale_shape_discrete(
    #name = "Sulfur dioxide\nconcentration (1994)",
    name = expression(SO[2] * " concentration (1994)"),
    labels = c(
      "40" = expression("40" * mu * "g"/m^3),
      "50" = expression("50" * mu * "g"/m^3),
      "60" = expression("60" * mu * "g"/m^3)
    )
  ) +
  scale_y_continuous(
    name = "Difference from the reference [p.p.]"
  ) +
  theme_classic(base_family = "Times") +
  theme(
    axis.title.x = element_blank()
  )
```